---
title: "Wine quality"
author: "Ryan Goding, George Jiao and Yujie Liu"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# The Setup

```{r load packages, message=FALSE}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(sandwich)
library(lmtest)
library(fec16)
library(MASS)
theme_set(theme_minimal())
knitr::opts_chunk$set(dpi = 300)
```

# Data Cleaning

```{r, message=FALSE}
df <- read.table("./winequality.csv", header = TRUE, sep = ",")
```

# Data Exploration

Scatter plot of each variable against quality

```{r, message=FALSE}
fixed_acid <- df %>% 
  ggplot() + 
  aes(x = fixed.acidity, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Fixed Acidity', 
    x        = 'Fixed Acidity', 
    y        = 'Quality'
  )

volatile_acid <- df %>% 
  ggplot() + 
  aes(x = volatile.acidity, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Volatile Acidity', 
    x        = 'Volatile Acidity', 
    y        = 'Quality'
  )

citric_acid <- df %>% 
  ggplot() + 
  aes(x = citric.acid, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Citric Acid', 
    x        = 'Citric Acid', 
    y        = 'Quality'
  )

residual_sugar <- df %>% 
  ggplot() + 
  aes(x = residual.sugar, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Residual Sugar', 
    x        = 'Residual Sugar', 
    y        = 'Quality'
  )

chlorides <- df %>% 
  ggplot() + 
  aes(x = chlorides, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Chlorides', 
    x        = 'Chlorides', 
    y        = 'Quality'
  )

free_sulfur_dioxide <- df %>% 
  ggplot() + 
  aes(x = free.sulfur.dioxide, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Free Sulfur Dioxide', 
    x        = 'Free Sulfur Dioxide', 
    y        = 'Quality'
  )

total_sulfur_dioxide <- df %>% 
  ggplot() + 
  aes(x = total.sulfur.dioxide, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Total Sulfur Dioxide', 
    x        = 'Total Sulfur Dioxide', 
    y        = 'Quality'
  )

density_ <- df %>% 
  ggplot() + 
  aes(x = density, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Density', 
    x        = 'Density', 
    y        = 'Quality'
  )

pH <- df %>% 
  ggplot() + 
  aes(x = pH, y = quality) + 
  geom_point() + 
  labs(
    title    = 'pH', 
    x        = 'pH', 
    y        = 'Quality'
  )

sulphates <- df %>% 
  ggplot() + 
  aes(x = sulphates, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Sulphates', 
    x        = 'Sulphates', 
    y        = 'Quality'
  )

alcohol <- df %>% 
  ggplot() + 
  aes(x = alcohol, y = quality) + 
  geom_point() + 
  labs(
    title    = 'Alcohol', 
    x        = 'Alcohol', 
    y        = 'Quality'
  )

fixed_acid
volatile_acid
citric_acid
residual_sugar
chlorides
free_sulfur_dioxide
total_sulfur_dioxide
density_
pH
sulphates
alcohol

```

There seems to have some slight negative correlation between quality and volatile acidity, quality and density. Slight positive correlation between quality and alcohol.

The difficulty of using linear regression is quality only takes value between 3 and 8.

Exploring the distribution of quality below:

```{r distriution of outcome data}
quality_histogram <- df %>% 
  ggplot() + 
  aes(x = quality) + 
  geom_histogram(bins = 30) + 
  labs(
    title    = 'Quality', 
    x        = 'Quality', 
    y        = 'Frequency'
  )

quality_histogram
```

We have to transform the quality variable in some way. Try logit transformation.

```{r transformation}
model <- polr(as.factor(quality) ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = df, Hess = TRUE)
```

